# Data sources
## Sources  

These three sets record vehicle accidents in States of New York from 2018 to 2019. Data are from US government public data. The three sets contain different attributes about each vehicle accident reported to NYS DMV.  

**Data of Vehicle Information**  
https://catalog.data.gov/dataset/motor-vehicle-crashes-vehicle-information-three-year-window

**Data of Individual Information**  
https://catalog.data.gov/dataset/motor-vehicle-crashes-individual-information-three-year-window

**Data of Case Information**  
https://catalog.data.gov/dataset/motor-vehicle-crashes-case-information-three-year-window


## Basic Information  

**Data of Vehicle Information**  
```{r,message = FALSE,echo = FALSE}
library(tidyverse)
vehicle_data_full <- read_csv("~/GR5702/Project/vehicle_original.csv")
```

```{r,message = FALSE,echo = FALSE}
dim <- dim(vehicle_data_full)
paste0("There are ",as.character(dim[1])," rows and ",as.character(dim[2])," variables in this dataset.")
print("Variables contains:")
colnames(vehicle_data_full)
```

**Data of Case Information**  
```{r,message = FALSE,echo = FALSE}
case_data_full <- read_csv("~/GR5702/Project/case_original.csv")
```

```{r,message = FALSE,echo = FALSE}
dim <- dim(case_data_full)
paste0("There are ",as.character(dim[1])," rows and ",as.character(dim[2])," variables in this dataset.")
print("Variables contains:")
colnames(case_data_full)
```

**Data of Individual Information**  
```{r,message = FALSE,echo = FALSE}
individual_data_full <- read_csv("~/GR5702/Project/individual_original.csv")
```

```{r,message = FALSE,echo = FALSE}
dim <- dim(individual_data_full)
paste0("There are ",as.character(dim[1])," rows and ",as.character(dim[2])," variables in this dataset.")
print("Variables contains:")
colnames(individual_data_full)
```

## Issues or Problems in Data  
Since the data are collected and published by the government, the data is relatively clean. Thus, the categorical inputs are all understandable and numerical inputs are mostly in reasonable domain.  
However, some variable have large amount of missing values, which cause some troubles in the later section, but the difficulty is manageable. Also, the three datasets cannot be merged together because they do not have a common key, but fortunately, each dataset have good amount of variables to explore.  

## Potential Questions  
To be more specific, our interest falls on several aspects: 

About Accident Cases:
1.	Distribution of the number of accidents across time (Date, time, weekdays)?
2.	Any environmental factors (weather condition, road surface, lighting condition) related to accidents?

About Drivers:
1.	Association between Injured severity and safety equipment?
2.	Association between Injured severity and seating position?

About Vehicle:
1.	Relationship between number of accidents and Vehicle year?
2.	Relationship between number of accidents and number of occupants?
3.	Relationship between number of accidents and number of vehicle cylinder?

The original data contains more than 800,000 samples each. For the convenience of later use, we did some random sampling. Before diving into the visualization work, we found some potential problems around the our data. Firstly, there are some misssing values in the data. Secondly, we need to select some variales that contribute the most to car crashes. Lastly, since most of our variables are categorical, we need to find better methods for visualization.