# Missing values
Since we have three datasets and they do not have a common key attribute for us to merge them together, we decided to visualize the missing values in three seperate graphs.  

## Vehicle
```{r,message = FALSE,echo = FALSE}
library(tidyverse)
library(naniar)
```

```{r,message = FALSE,echo = FALSE}
df1 <- read_csv("~/GR5702/Project/Vehicle.csv")
```

```{r,echo = FALSE}
label.na <- c("Unknown","Not Applicable","UNKNOWN VEHICLE","Not Entered","None")
for (i in 1:ncol(df1)) {
  for (j in 1:nrow(df1)) {
    if(df1[j,i] %in% label.na ){
      df1[j,i] <- NA
      }
  }
}
```

```{r,echo = FALSE}
gg_miss_var(df1, show_pct = TRUE)
```

In this dataset, "Type/Axles of Truck or Bus" has all missing values in the column. "Event Type" and "Contributing Factor 1&2 Descriptions" also have many missing values. The other variables contain far less missing values.  

## Case
```{r,message = FALSE,echo = FALSE}
df2 <- read_csv("~/GR5702/Project/Case.csv")
```

```{r,echo = FALSE}
label.na <- c("Unknown","Not Applicable","UNKNOWN VEHICLE","Not Entered","None")
for (i in 1:ncol(df2)) {
  for (j in 1:nrow(df2)) {
    if(df2[j,i] %in% label.na ){
      df2[j,i] <- NA
      }
  }
}
```

```{r,echo = FALSE}
gg_miss_var(df2, show_pct = TRUE)
```

In this dataset, We had 18 variables in total and we dropped 2 of them. The majority of "DOT.Reference.Marker.Location" are missing values and the "Pedestrian.Bicyclist.Action" contains a lot of "Not Applicable" values which cannot offer the information we want. 

After transformation and cleaning of the dataset, we got the final data for visualization use. 

## Individual
```{r,message = FALSE,echo = FALSE}
df3 <- read_csv("~/GR5702/Project/Individual.csv")
```

```{r,echo = FALSE}
label.na <- c("Unknown","Not Applicable","UNKNOWN VEHICLE","Not Entered","None")
for (i in 1:ncol(df3)) {
  for (j in 1:nrow(df3)) {
    if(df3[j,i] %in% label.na ){
      df3[j,i] <- NA
      }
  }
}
```

```{r,echo = FALSE}
gg_miss_var(df3, show_pct = TRUE)
```

In this dataset, we have 15 variables in total. "Transported_by", "Injury Location", "Victim Status" and "Injury Description" have high percent of missing data within the variable.  